### 2024-04-19 paper/paper.qmd
new change

---
title: "How do domestic GDP growth, global economic conditions, and financial crises impact the export growth rate in different countries?"
author: "Mohammed Yusuf Shaikh"
thanks: "Code and data are available at: https://github.com/Yusuf365/Data-Analysis-Regression.git"
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
### Load Data and Packages

install.packages("readr")
install.packages("MASS")
install.packages("rstanarm")
install.packages("bayesplot")
install.packages("modelsummary")
install.packages("AER")
install.packages("broom")
install.packages("stargazer")
library(stargazer)
library(readr)
library(MASS)
library(rstanarm)
library(bayesplot)
library(modelsummary)
library(ggplot2)
library(broom)
library(knitr)
library(tidyverse)
library(AER)
library(opendatatoronto)
library(dplyr)
library(readr)
library(janitor)
library(ggplot2)
library(readr)
```

```{r}
#| include: false
#| warning: false
#| message: false
data <-
  read_csv(
    here::here("/cloud/project//data/analysis_data/analysis_data.csv")
  )
```

# Introduction {#sec-intro}

# Data {#sec-data}

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-raw_data
#| tbl-cap: Raw Data Preview

kable(head(data))
```

# Model {#sec-model}

Data used in this paper was cleaned, processed, modeled and tested with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx], `viridis` [@citeViridis], `lubridate` [@citeLubridate].

# Results

# Discussion

# Conclusion

# References